|

### 2022-04-15 article/article.Rmd
proofreading

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: Legislature Size; Legislative Bargaining; Local Public Goods Provision; Regression Discontinuity Design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Legislative institutions are crucial for citizen welfare and public service
provision in democratic countries. Historically, the division of power and the
birth of representative governments placed legislatures at the center of the
political stage. In modern democracies, most countries have assemblies at all
levels of government. These institutions regulate the quantity and quality of
legislation, check and veto other political powers, and influence policymaking.
Moreover, legislatures also determine taxation, government spending and oversee
public programs \citep{auriol2012optimal, north1989constitutions,
poulsen2018political, stasavage2011states, Weingastetal1981}.

@@

In this paper, we propose a new theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
First, in mayor-council municipalities, mayors are the main responsible for
public service decisions and have the power to boost public investment similar
to presidents and governors. Second, partisanship plays a key role in city
politics, even in countries in which ideological attachment is relatively weak.
Party attachment structures political views \citep{samuels2014power}, and
local-level elites usually belong to different factions and parties. They fight
for power in the municipality by controlling various parties and forming
distinct pre-electoral coalitions.\footnote{Elite political competition in the
local level has also been linked to violence and conflict in Brazil and
elsewhere \citep{hoelscher2015politics, mignozzetti2018predatory}.} In many
cities, family ties may be substitutes for the ideological component, but
influential local leaders usually sort themselves into different political
groups \citep{ferraz2020political, nunes1977coronelismo}.

This paper incorporates partisanship and mayoral first-move advantage to a
\citet{baron1989bargaining} bargaining model. Mayors make the initial policy
proposal, which consists of investments in public goods provision, rents for
personal gains, and a vector of side-transfers for councilors. If mayors fail
to secure majoritarian support from councilors, a reversal stage occurs, in
which the decision leaves the mayor's office and councilors propose the public
investments. In this mayor-council bargaining process, costs come from the
side-transfers required to secure a majority among councilors. We show that in
a model without partisan concerns, bargaining costs always increase with larger
city councils, which in turn lower public goods provision and rents. However,
partisanship decreases bargaining costs when the chances of electing councilors
aligned with the mayor are sufficiently high. Lower bargaining costs release
funds for mayors to invest in public goods provision. In this case, a larger
city council may have more partisan councilors aligned with the mayor, reducing
the mayor's bargaining costs and increasing the equilibrium public goods
provision.

@@

To provide further evidence for our mechanisms, we surveyed 174 former
councilors who served during the 2005-2008 term, the period we analyse in this
paper. They confirm that mayors indeed use bureaucratic appointments and
political favors to secure legislative support. As our formal model shows,
bargaining with councilors requires offering transfers to the representatives.
By asking councilors what they would like to provide to voters, we find that
councilors mainly demand job appointments, as these increase their reelection
prospects and improve municipal welfare. These findings are in line with
quantitative results from a novel dataset we built with 346,553 laws passed in
63 municipalities within ten thousand inhabitants away from the cutoffs. We see
that legislation in municipalities both on the left- and on the right-hand side
of the cutoffs mention public goods provision. However, municipalities with
larger city councils have 15 percent more proposals about public goods
provision, which is consistent with our argument.

Finally, we show that the increase in legislature size significantly impacts
municipal education and health care. Increasing the city council size by one
legislator lowers infant mortality by 2.01 children per one thousand infants
born and reduces postneonatal mortality by 0.90 children per one thousand
infants that survived their first 28 days. Larger councils also increase
enrollment by 2.58 children in elementary education classrooms without reducing
school quality. These services are crucial for citizen welfare and result from
eased bargaining between the mayor and the city council.

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, who
extended the bilateral bargaining game proposed by
\citet{rubinstein1982perfect} to more than two players. Several papers followed
\citet{baron1989bargaining} and included endogenous recognition power, the size
of the required majority, and partisanship in the model
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}.\footnote{See
\citet{eraslan2019legislative} for a comprehensive review of multilateral
bargaining and \citet{ray2015coalition} for a review on coalition formation.}
In this paper, we extend the \citet{baron1989bargaining} model under closed
rules and add the mayor's first move and political partisanship. We consider
partisanship as a zero-sum effect that is positive for politicians aligned with
the mayor and negative otherwise. This paper investigates the often overlooked
effect of legislature size on the bargaining equilibrium strategies.

@@

Despite the importance of legislatures, the executive still dominates the
public debate on public service provision in most mayor-council (or
presidential) systems. For instance, while in 2021 the New York Times cited
Mayor Bill De Blasio 2,084 times, it mentioned the 49 seating city councilors
only 440 times in the same year. When we expand the search for the past five
years, we see that De Blasio was mentioned 8,978 while the 49 seating city
councilors were mentioned 2,046 times.\footnote{We computed these figures using
the Dow Jones Factiva Database in November 2021.} On average, the mayor was
cited four times more often than all the city councilors combined. This
executive dominance has implications for policymaking, as coalitions and
affinity with the executive are going to be crucial for service provision
\citep{buenodemesquita2005logic}.

@@

Solving the game by backward induction requires us to start with the
councilors' strategies. However, to conserve space, assume that we solved the
game for the councilors and found the equilibrium bargaining costs $C_G(N)$
when there are $N$ city councilors. We now derive the optimal rents ($r$) and
public goods provision ($g$) proposed by the mayor. The mayor benefits from
public goods provision, as it increases the chances of her reelection. However,
she prefers to invest as minimum as necessary in public goods and extract the
remaining resources as political rents. The mayor's objective is to maximize
her expected utility, subject to the municipal budget constraint.

@@

\begin{proposition}\label{prop1}
Public goods provision increases with legislature size when the bargaining costs decrease with legislature size.
\end{proposition}

@@

Also, the threshold $\underline{\gamma}$ that captures the probability of
electing a mayor-aligned policitian decreases along with the political value of
policy $p$ and the size of the legislature $N$. This means that intensive
partisanship and larger legislatures make it easier to satisfy the electoral
threshold.

@@

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=H\arabic*.]
  \item Bargaining costs decrease with legislature size when the chances of electing a government-aligned legislator are sufficiently high.
  \item Public goods provision increases if bargaining costs decrease with legislature size.
\end{enumerate}

@@

Until 2004, there was no specific regulation on the size of city councils. The
1988 Brazilian Constitution set only broad guidelines about the number of
council seats in each municipality, thus allowing cities to determine how many
legislators they should have in their local chambers. This led to severe
imbalances in the proportion of representatives per citizen across the country.
For instance, the Nova Russas municipality had 30,009 inhabitants and 21
councilors, accounting for one legislator per 1,429 citizens. In contrast, the
city of Sorocaba had only 15 councilors for 528,735 inhabitants, that is, one
legislator per 35,249 citizens. In 2003, this malapportionment problem gained
national visibility when Mira Estrela, a small municipality with only 2,651
residents, reduced its city council from 11 to 9 seats because of fiscal
considerations. The change motivated a legal dispute that reached the Brazilian
Supreme Court. The Court favored the city's decision and ruled that the size of
every local council should correspond to precise population thresholds.

@@

Brazil also adopts a unique formula to allocate seats, which combines the Hare
quota and the D'Hondt method and favors larger parties
\citep{nicolau2015sistemas}. This poses a major obstacle for small parties to
reach the electoral quorum, so they often join party coalitions to improve
their chances of winning legislative seats \citep[p. 68]{Ames2001}. In Brazil,
mayors are usually members of the largest coalitions
\citep{poulsen2018political}, thus there is an incentive for candidates to side
with the mayor before and after the election.

@@

Third, we also collected 346,553 bills from cities within ten thousand
inhabitants from the 2004 population thresholds. We selected 63 out of 202
municipalities whose voting records were available online. We evaluate proposal
types, discriminating between local public goods, oversight, and other
legislative activity. Examples of the local public goods are investing in a
public health clinic, requesting school staff to accommodate an extra child,
renovating sewage, and fixing potholes on a given street.\footnote{Our theory
does not differentiate between clientelism and broad service provision, as it
is hard to disentangle differences in the local level legislation. A voter may
request a particular service, such as fixing a pothole. Attending this request
can be interpreted as a clientelistic practice. Still, the provision of the
service generates non-excludable benefits to the adjacent neighborhood.}
Oversight bills are accountability and transparency mechanisms meant to keep
public services provision in check. Other legislation comprises bills that do
not fit in the previous categories, such as changes in street names, motions to
honor citizens or groups, and legislative and internal city council procedures.
We provide further information about the legislation dataset and the city
council survey in sections A.13 and A.14 of the Supplementary Materials.

@@

Together, this evidence provides a clear picture of local-level politics in
Brazil, from the political mechanism to the municipal welfare consequences of
the council size ruling. Summary statistics and variable descriptions follow in
section A.2 of the Supplementary Materials.

@@

@@

Column (1) shows a 91\% chance of the extra legislator belonging to the mayoral
coalition. The model hypothesis stated that the probability of electing a city
councilor aligned with the mayor is sufficiently high. However, sufficiently
high can be ambiguous, and it depends on the intensity of the political
disagreement and the current legislature size. In any case, 91\% chance is a
considerably large quantity, and the extra city councilor is going to the
mayoral coalition nine in every ten legislature size increases. This
strengthens the political position of the mayor and diminishes bargaining
costs.\footnote{The effect of city council size on the number of mayor-aligned
politicians stems from the rounding rule adopted by the proportional
representation system in Brazil. The rounding rule favors the coalitions with
more votes. Most of the time, the coalition of the elected mayor is the one
that receives more votes.} Column (2) demonstrates that adding one councilor
increases appointed employees by around 104 extra employees per additional city
councilor. When negotiating with city councilors, these employees are a great
bargaining chip, and they are also associated with better service provision
\citep{toral2019benefits}. As expected, legislature size has a null effect on
the number of career bureaucrats (Column 3) and councilors' staff members
(Column 4).

@@

In terms of education, with the addition of a city councilor, enrollment in
elementary schools increased by 2.58 children on average per school. This
represents a change of around 0.20 standard deviations. Besides the influence
of the mayoral support, enrollment can also directly impact the councilor's
representation: Councilors can pressure the bureaucracy to accommodate a few
extra students. With support from the mayor, they can help voters get access to
schools, improving the welfare of voters. Education quality remains unchanged,
but with more students in the classroom, it could potentially decrease. As
expected, the placebo regressions are insignificant.

@@

@@

In this paper, we show that the provision of education and health care in
Brazil improves with larger legislatures. This is because the chances of
electing a city councilor aligned with the mayor are 91 percent, thanks to the
rule of distributing seats adopted in the country. In the Supplemental
Materials, we also run regressions in placebo cutoffs between the actual
cutoffs, showing that our estimates are consistently null when we are indeed
supposed to find insignificant results.

@@

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, and \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains uncertain
how other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about how local
legislatures impact public service delivery. These questions are crucial for
institution design and welfare in developing democracies.

>>>>

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed 174 former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: legislature size; legislative bargaining; local public goods provision; regression discontinuity design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Legislative institutions are crucial for citizen welfare and public service
provision in democratic countries. Historically, the separation of powers and
the birth of representative governments placed legislatures at the center of
the political stage. In modern democracies, most countries have assemblies at
all levels of government. These institutions regulate the quantity and quality
of legislation, check and veto other political powers, and influence
policymaking. Moreover, legislatures also determine taxation, government
spending and oversee public programs \citep{auriol2012optimal,
north1989constitutions, poulsen2018political, stasavage2011states,
Weingastetal1981}.

@@

In this paper, we propose a theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
First, in mayor-council municipalities, mayors are the main responsible for
public service decisions, and like presidents and governors, they can boost
local public investments by altering the government budget. Second,
partisanship plays a key role in city politics, even in countries in which
ideological attachment is relatively weak. Party attachment structures
political views \citep{samuels2014power}, and local-level elites usually belong
to different factions and parties. They fight for power in the municipality by
controlling various parties and forming distinct pre-electoral
coalitions.\footnote{Elite political competition in the local level has also
been linked to violence and conflict in Brazil and elsewhere
\citep{hoelscher2015politics, mignozzetti2018predatory}.} In many cities,
family ties may be substitutes for the ideological component, but influential
local leaders usually sort themselves into different political groups
\citep{ferraz2020political, nunes1977coronelismo}.

This paper incorporates partisanship and mayoral first-move advantage to a
\citet{baron1989bargaining} bargaining model. Mayors make the initial policy
proposal, which consists of investments in public goods provision, rents for
personal gains, and a vector of side-transfers for councilors. If mayors fail
to secure majoritarian support from councilors, a reversal stage occurs, in
which the decision leaves the mayor's office and councilors propose the public
investments. In this mayor-council bargaining process, costs come from the
side-transfers required to secure a majority among councilors. We show that in
a model without partisan concerns, bargaining costs always increase with larger
city councils, which in turn lower public goods provision and rents. However,
partisanship decreases bargaining costs when the chances of electing
government-aligned councilors are sufficiently high. Lower bargaining costs
release funds for mayors to invest in public goods provision. In this case, a
larger city council may add more members to the government coalition, thus
reducing the mayor's bargaining costs and increasing the equilibrium public
goods provision.

@@

To provide further evidence for our mechanisms, we surveyed 174 former
councilors who served during the 2005-2008 term, the period we analyse in this
paper. They confirm that mayors indeed use bureaucratic appointments and
political favors to secure legislative support. As our formal model shows,
bargaining with councilors requires offering transfers to the representatives.
By asking councilors what they would like to provide to voters, we find that
councilors mainly demand job appointments, which are regarded as an effective
means to boost their reelection prospects and improve municipal welfare. These
findings are in line with quantitative results from a novel dataset we built
with 346,553 laws passed in 63 municipalities within ten thousand inhabitants
away from the cutoffs. While all municipalities mention public goods in their
legislation, those with larger city councils have 15 percent more proposals
about public service delivery, which is consistent with our argument.

Finally, we show that the increase in legislature size significantly impacts
municipal education and health care. Increasing the city council size by one
legislator lowers infant mortality by 2.01 children per one thousand infants
born and reduces postneonatal mortality by 0.90 children per one thousand
infants who survived their first 28 days. Larger councils also increase
enrollment by 2.58 children in elementary education classrooms without reducing
school quality. These services are crucial for citizen welfare and result from
eased bargaining between the mayor and the city council.

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, who
extended the bilateral bargaining game proposed by
\citet{rubinstein1982perfect} to more than two players. Several papers followed
\citet{baron1989bargaining} and included endogenous recognition power, the size
of the required majority, and partisanship in the model
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}.\footnote{See
\citet{eraslan2019legislative} for a comprehensive review of multilateral
bargaining and \citet{ray2015coalition} for a review on coalition formation.}
In this article, we extend the \citet{baron1989bargaining} model under closed
rules and add the mayor's first move and political partisanship. We consider
partisanship as a zero-sum effect that is positive for politicians aligned with
the mayor and negative otherwise. We investigate the often overlooked effect of
legislature size on the bargaining equilibrium strategies.

@@

Despite the importance of legislatures, the executive still dominates the
public debate on public service provision in most mayor-council (or
presidential) systems. For instance, while in 2021 the New York Times cited
Mayor Bill De Blasio 2,084 times, it mentioned the 49 seating city councilors
only 440 times in the same year. When we expand the search for the past five
years, we see that De Blasio was mentioned 8,978 times while the 49 seating
city councilors were mentioned only 2,046 times.\footnote{We computed these
figures using the Dow Jones Factiva Database in November 2021.} On average, the
mayor was cited four times more often than all the city councilors combined.
This executive dominance has implications for policymaking, as coalitions and
affinity with the executive are going to be crucial for service provision
\citep{buenodemesquita2005logic}.

@@

Solving the game by backward induction requires us to start with the
councilors' strategies. To conserve space, assume that we solved the
game for the councilors and found the equilibrium bargaining costs $C_G(N)$
when there are $N$ city councilors. We now derive the optimal rents ($r$) and
public goods provision ($g$) proposed by the mayor. The mayor benefits from
public goods provision, as it increases the chances of her reelection. However,
she prefers to invest as minimum as necessary in public goods and extract the
remaining resources as political rents. The mayor's objective is to maximize
her expected utility, subject to the municipal budget constraint.

@@

\begin{proposition}\label{prop1}
Public goods provision increases with legislature size if bargaining costs decrease with legislature size.
\end{proposition}

@@

Also, the threshold $\underline{\gamma}$ that captures the probability of
electing a government supporter decreases along with the political value of
policy $p$ and the size of the legislature $N$. This means that intensive
partisanship and larger legislatures make it easier to satisfy the electoral
threshold.

@@

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=H\arabic*.]
  \item Bargaining costs decrease with legislature size when the chances of electing a government partisan are sufficiently high.
  \item Public goods provision increases if bargaining costs decrease with legislature size.
\end{enumerate}

@@

Until 2004, there was no specific regulation on the size of city councils. The
1988 Brazilian Constitution set only broad guidelines about the number of
council seats in each municipality, allowing cities to determine how many
legislators their local chambers should have. This led to severe imbalances in
the representation ratio across the country. For instance, the Nova Russas
municipality had 30,009 inhabitants and 21 councilors, or one legislator per
1,429 citizens. In contrast, the city of Sorocaba had only 15 councilors for
528,735 inhabitants, that is, one legislator per 35,249 citizens. In 2003, this
malapportionment problem gained national visibility when Mira Estrela, a small
municipality with only 2,651 residents, reduced its city council from 11 to 9
seats because of fiscal considerations. The change motivated a legal dispute
that reached the Brazilian Supreme Court. The Court favored the city's decision
and ruled that the size of every local council should correspond to precise
population thresholds.

@@

Brazil also adopts a unique formula to allocate seats, which combines the Hare
quota and the D'Hondt method and strongly favors larger parties
\citep{nicolau2015sistemas}. This poses a major obstacle for small parties to
reach the electoral quorum, so they often join party coalitions to improve
their chances of winning legislative seats \citep[p. 68]{Ames2001}. In Brazil,
mayors are usually members of the largest coalitions
\citep{poulsen2018political}, thus there is an incentive for candidates to side
with the mayor before and after the election.

@@

Third, we collected 346,553 bills from cities within ten thousand inhabitants
from the 2004 population thresholds. We selected 63 out of 202 municipalities
whose voting records were available online. We evaluated proposal types,
discriminating between local public goods, oversight, and other legislative
activity. Examples of the local public goods are investing in a public health
clinic, requesting school staff to accommodate an extra child, renovating
sewage, and fixing potholes on a given street.\footnote{Our theory does not
differentiate between clientelism and broad service provision, as it is hard to
disentangle differences in the local level legislation. A voter may request a
particular service, such as fixing a pothole. Attending this request can be
interpreted as a clientelistic practice. Still, the provision of the service
generates non-excludable benefits to the adjacent neighborhood.} Oversight
bills are accountability and transparency mechanisms meant to keep public
services provision in check. Other legislation comprises bills that do not fit
in the previous categories, such as changes in street names, motions to honor
citizens or groups, and legislative and internal city council procedures. We
provide further information about the legislation dataset and the city council
survey in sections A.13 and A.14 of the Supplementary Materials.

@@

Together, this evidence provides a clear picture of local-level politics in
Brazil, from the political mechanism to the municipal welfare consequences of
the council size ruling. Summary statistics and variable descriptions are
available in section A.2 of the Supplementary Materials.

@@

\vspace{.4cm}

@@

Column (1) shows a 91\% chance of the extra legislator belonging to the mayoral
coalition. The model hypothesis stated that the probability of electing a city
councilor who is aligned with the mayor is sufficiently high. However,
sufficiently high can be ambiguous, and it depends on the intensity of the
political disagreement and the current legislature size. In any case, 91\%
chance is a considerably large quantity, and the extra city councilor is going
to the mayoral coalition nine in every ten legislature size increases. This
strengthens the political position of the mayor and diminishes bargaining
costs.\footnote{The effect of city council size on the number of mayor-aligned
politicians stems from the rounding rule adopted by the proportional
representation system in Brazil. The rounding rule favors the coalitions with
more votes. Most of the time, the coalition of the elected mayor is the one
that receives more votes.} Column (2) demonstrates that adding one councilor
increases appointed employees by around 104 extra employees per additional city
councilor. When negotiating with city councilors, these employees are a great
bargaining chip, and they are also associated with better service provision
\citep{toral2019benefits}. As expected, legislature size has a null effect on
the number of career bureaucrats (Column 3) and councilors' staff members
(Column 4).

@@

@@

In terms of education, with the addition of a city councilor, enrollment in
elementary schools increased by 2.58 children on average per school. This
represents a change of around 0.20 standard deviations. Besides the influence
of the mayoral support, enrollment can also directly impact the councilor's
representation: Councilors can pressure the bureaucracy to accommodate a few
extra students. With support from the mayor, they can help voters get access to
schools, improving the welfare of voters. Education quality remains unchanged,
but with more students in the classroom, it could potentially decrease. Again,
the placebo regressions are insignificant.

@@

In this paper, we show that the provision of education and health care in
Brazil improves with larger legislatures. This is because the chances of
electing a government-aligned candidate are 91 percent thanks to the rule of
distributing seats adopted in the country. In the Supplemental Materials, we
also run regressions in placebo cutoffs between the actual cutoffs, showing
that our estimates are consistently null when we are indeed supposed to find
insignificant results.

@@

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, and \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains unclear how
other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about the impact of
local and national legislatures impact public services. These questions are
crucial for institution design and welfare in developing democracies.
